---
title: "HovanesKeseyan_Lab10"
author: "Hovanes Keseyan"
date: "November 29, 2016"
output: pdf_document
---

## Testing for Outliers

```{r }
x <- c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, 2.0, -5.7, -7.9, -4.9, 28.7, 4.9)
```

## 1 : Calculating quartiles

```{r }
#xs <- sort(x)

# first quartile
q1 <- quantile(x, 0.25)
q1

# third quantile
q3 <- quantile(x, 0.75)
q3

# interquartile range
iqr <- q3 - q1
iqr
```

## 2 : quartiles function

```{r }
quartiles <- function(x){
  output <- vector("list", 3)
  names(output) <- c("First quartile", "Third quartile", "Inter-quartile range")
  
  q1 <- quantile(x, 0.25)
  output[1] <- q1
  q3 <- quantile(x, 0.75)
  output[2] <- q3
  iqr <- q3 - q1
  output[3] <- iqr
  
  return(output)
}

quartiles(x)
```

## 3 : Tukey's rule outliers

```{r }
tukey.low <- q1 - 1.5*iqr
tukey.low

tukey.high <- q3 + 1.5*iqr
tukey.high

sort(x)
```

The only outlier is 28.7 (the 12th entry), because it is the only value in the vector that is lower than the low end or higher than the high end of the values calculated by Tukey's rule.

## 4 & 5 : Tukey Outlier functions

```{r }
tukey.outlier <- function(x){
  n <- length(x)
  outliers <- logical(n)
  q <- quartiles(x)
  tukey.low <- q[[1]] - 1.5*q[[3]]
  tukey.high <- q[[3]] + 1.5*q[[3]]
  # if value is outlier, return TRUE for corresponding entry in the outliers
  outliers <- ifelse(x < tukey.low | x > tukey.high, T, F)
  return(outliers)
}

tukey.outlier(x)

test.tukey.outlier <- function(x){
  # initialize vector of manual outliers found by observation
  mo <- logical(length(x))
  # we know that the only outlier is the 12th entry (28.7)
  mo[12] <- T
  to <- tukey.outlier(x)
  # create a  vector that stores FALSE for each entry where function fails
  test.results <- ifelse(to == mo, T, F)
  # the test works if and only if all values of the vector are TRUE
  response <- ifelse(all(test.results),T, c("The function fails the test."))
  return(response)
}

test.tukey.outlier(x)
```

## 6 : -x

```{r }
q <- quartiles(-x)
tukey.low <- q[[1]] - 1.5*q[[3]]
tukey.low
tukey.high <- q[[3]] + 1.5*q[[3]]
tukey.high
sort(-x)
-x
```

Again, the 12th entry (this time, -28.7) is the only outlier.

## 7 : 100*x

```{r }
q <- quartiles(100*x)
tukey.low <- q[[1]] - 1.5*q[[3]]
tukey.low
tukey.high <- q[[3]] + 1.5*q[[3]]
tukey.high
sort(100*x)
100*x
```

Again, the 12th entry (this time, 2870) is the only outlier.

## 8 & 9 : Test Tukey outlier on new cases

```{r }
tukey.outlier(-x)
test.tukey.outlier(-x)

tukey.outlier(100*x)
test.tukey.outlier(100*x)
```

The function does not need to be modified, since in each case we had the same set of entries (only entry 12) as the outliers. The function passes the test each time.

## 10 : 

```{r }
y <- c(11.0, 14.0, 3.5, 52.5, 21.5, 12.7, 16.7, 11.7, 10.8, -9.2, 12.3, 13.8, 11.1)
q <- quartiles(y)
tukey.low <- q[[1]] - 1.5*q[[3]]
tukey.low
tukey.high <- q[[3]] + 1.5*q[[3]]
tukey.high
sort(y)
y
```

We have the 10th and 3rd entries (-9.2 and 3.5) as low outliers, and every other entry as high outliers. We need to modify our test.tukey.outlier function to account for all of the values in the data set as outliers.

```{r }
test.tukey.outlier <- function(x){
  # initialize vector of manual outliers found by observation
  mo <- logical(length(x))
  # we know that the only outlier is the 12th entry (28.7)
  mo[1:13] <- T
  to <- tukey.outlier(x)
  # create a  vector that stores FALSE for each entry where function fails
  test.results <- ifelse(to == mo, T, F)
  # the test works if and only if all values of the vector are TRUE
  response <- ifelse(all(test.results),T, c("The function fails the test."))
  return(response)
}

tukey.outlier(y)
test.tukey.outlier(y)
```

The function still passes the test after modifying the test to reflect what we observed for this new data set.
