---
title: "HovanesKeseyan_Homework6"
author: "Hovanes Keseyan"
date: "November 5, 2016"
output: pdf_document
---

## Part I : Gambler's Ruin

## 1 : Betting on a fair game

```{r }
# starting bankroll
sbr <- 1000
# bet amount
b <- 100
# max iterations
n <- 5000
# probability of winning a hand
p <- 0.5

```

The following function takes the starting bankroll, the bet amount, the maximum number of iterations, and the probability of winning a hand, and returns a boolan value informing whether or not the player has busted. A return value of TRUE indicates the player has busted.

```{r }
gamble <- function(br=sbr, b=b, n=n, p=p) {
  i <- 0
  while(br > 0 & i < n){
    wl <- sample(c(1, -1), size=1, prob = c(p, 1-p))   # win/loss outcome
    br <- br + wl*b
    i <- i+1
  }
  if(i >= n) busted <- FALSE
  else busted <- TRUE
  return(busted)
}
```

## 1a : Probability of having busted by 100th bet

```{r }
# run the gamble function a large number of times and find probability
p.b.100 <- sum(replicate(1000,gamble(sbr,b,100, 0.5)))/1000
p.b.100
```

## 1b : Probability of having busted by 500th bet

```{r }
# run the gamble function a large number of times and find probability
p.b.500 <- sum(replicate(1000,gamble(sbr,b,500, 0.5)))/1000
p.b.500
```

## 1c : Mean time until bust, given busting within 5000 hands

The following function runs the gamble() function, but instead returns the number of hands played before busting. If the player did not bust by the time the maximum number of hands (n) is played, it repeats the function until the condition is satisfied.

```{r }
gamble.hands <- function(br=sbr, b=b, n=n, p=p) {
  i <- 0
  while(br > 0 & i < n){
    wl <- sample(c(1, -1), size=1, prob = c(p, 1-p))   # win/loss outcome
    br <- br + wl*b
    i <- i+1
  }
  return(i)
}
# run the gamble.hands function a large number of times and take the mean
v1c <- replicate(1000,gamble.hands(sbr,b, 5000, 0.5))
mean(v1c)

v1c <- replicate(1,gamble.hands(sbr,b, 5000, 0.5))
mean(v1c)

v1c <- replicate(10,gamble.hands(sbr,b, 5000, 0.5))
mean(v1c)

v1c <- replicate(3000,gamble.hands(sbr,b, 5000, 0.5))
mean(v1c)

```

## 1d : Mean and variance of bankroll after 100 hands

The following function runs the gamble() function, but instead returns the current bankroll after n bets.

```{r }
gamble.br <- function(br=sbr, b=b, n=n, p=p) {
  i <- 0
  while(br > 0 & i < n){
    wl <- sample(c(1, -1), size=1, prob = c(p, 1-p))   # win/loss outcome
    br <- br + wl*b
    i <- i+1
  }
  return(br)
}
# run the gamble.br function a large number of times and take the mean and variance
v1d <- replicate(1000,gamble.br(sbr,b, 100, 0.5))
mean(v1d)
var(v1d)
```

## 1e : Mean and variance of bankroll after 500 hands

```{r }
# run the gamble.br function a large number of times and take the mean and variance
v1e <- replicate(1000,gamble.br(sbr,b, 500, 0.5))
mean(v1e)
var(v1e)
```

## 2 : Betting on black

## 2a : Probability of having busted by 500th bet

```{r }
# run the gamble function a large number of times and find probability
p.b.100 <- sum(replicate(1000,gamble(sbr,b,100, 18/38)))/1000
p.b.100
```

## 2b : Probability of having busted by 500th bet

```{r }
# run the gamble function a large number of times and find probability
p.b.500 <- sum(replicate(1000,gamble(sbr,b,500, 18/38)))/1000
p.b.500
```

## 2c : Mean time until bust, given busting within 5000 hands

```{r }
# run the gamble.hands function a large number of times and take the mean
v2c <- replicate(1000,gamble.hands(sbr,b, 5000, 18/38))
mean(v2c)
```

## 2d : Mean and variance of bankroll after 100 hands

```{r }
# run the gamble.br function a large number of times and take the mean and variance
v2d <- replicate(1000,gamble.br(sbr,b, 100, 18/38))
mean(v2d)
var(v2d)
```

## 2e : Mean and variance of bankroll after 500 hands

```{r }
# run the gamble.br function a large number of times and take the mean and variance
v2e <- replicate(1000,gamble.br(sbr,b, 500, 18/38))
mean(v2e)
var(v2e)
```

## 3 : Bootstrap

```{r }
bootstrap <- function(v) {
  newv <- v - mean(v) + sbr
  n <- 1000
  bstrap <- double(n)
  for (i in 1:n) {
    newsample <- sample(newv,length(v),replace = T)
    bstrap[i] <- mean(newsample)
  }
  return(bstrap)
}
var(bootstrap(v1d))
var(bootstrap(v1e))
var(bootstrap(v2d))
var(bootstrap(v2e))
```

## Part II : Elo Ratings

## 4 : Setup

```{r }
# true vector of ratings
tv <- integer(13)
for (i in 1:13) {
  tv[i] <- sample(seq(-2,2),1)
}
tv

# vector of current ratings
cv <- tv

# matrix of ratings over time
rm <- matrix(NA, nrow = 13, ncol = 10000)
rm[,1] <- cv
```

## 5 : Function to simulate a game

```{r }
game <- function(i, j, tv) {
  p.i <- 1/(1+exp(tv[j]-tv[i]))    # probability that player i will win
  i.win <- rbinom(1,1,p.i)           # return 1 if i wins, 0 if j wins
  return(i.win)
}
```

## 6 : Function to update ratings

```{r }
rating.update <- function(i, j, k, tv, cv) {
  i.win <- game(i, j, tv)
  p.i <- 1/(1+exp(tv[j]-tv[i]))
  if(i.win==1) {
    cv[i] <- cv[i] + k*(1-p.i)
    cv[j] <- cv[j] - k*(1-p.i)
  }
  else {
    cv[i] <- cv[i] - k*p.i
    cv[j] <- cv[j] + k*p.i
  }
  return(cv)
}
```

## 7 : Function to simulate game between random players

```{r }
random.match <- function(k, tv, cv) {
  # select 2 players to be matched against each other
  players <- sample(1:length(tv),2,replace = FALSE)
  # run the rating.update function
  cv <- rating.update(players[1], players[2], k, tv, cv)
  return(cv)
}
```

## 8 : Function to simulate tournament

```{r }
tournament <- function(n=10000, k, tv, cv){
  rm <- matrix(NA, nrow = 13, ncol = 10000)
  rm[,1] <- cv
  for (iter in 2:n) {
    rm[,iter] <- random.match(k, tv, cv)
  }
  return(rm)
}
```

## 9 : Simulate a tournament

```{r }
rm <- tournament(n=10000, k=0.01, tv, cv)

# create a matrix tracking the best rating over time
best.p <- apply(rm, 2, max)
# create a matrix tracking the worst rating over time
worst.p <- apply(rm, 2, min)

plot(best.p,ty="l",col="green",main="k=0.01",ylab="Rating",xlab="Match",ylim=c(-3,3))
lines(worst.p, col="red")
```

The best and worst rating values seem to converge to 2 and -2, respectively, which is what we observed from the true ratings.

## 10 : More simulations

```{r }
rm <- tournament(n=10000, k=0.03, tv, cv)
best.p <- apply(rm, 2, max)
worst.p <- apply(rm, 2, min)
plot(best.p,ty="l", col="green", main="k=0.03",ylab="Rating",xlab="Match",ylim=c(-3,3))
lines(worst.p, col="red")

rm <- tournament(n=10000, k=0.06, tv, cv)
best.p <- apply(rm, 2, max)
worst.p <- apply(rm, 2, min)
plot(best.p,ty="l", col="green", main="k=0.06",ylab="Rating",xlab="Match",ylim=c(-3,3))
lines(worst.p, col="red")

rm <- tournament(n=10000, k=0.1, tv, cv)
best.p <- apply(rm, 2, max)
worst.p <- apply(rm, 2, min)
plot(best.p,ty="l", col="green", main="k=0.1",ylab="Rating",xlab="Match",ylim=c(-3,3))
lines(worst.p, col="red")

rm <- tournament(n=10000, k=0.3, tv, cv)
best.p <- apply(rm, 2, max)
worst.p <- apply(rm, 2, min)
plot(best.p,ty="l", col="green", main="k=0.3",ylab="Rating",xlab="Match",ylim=c(-3,3))
lines(worst.p, col="red")

rm <- tournament(n=10000, k=0.6, tv, cv)
best.p <- apply(rm, 2, max)
worst.p <- apply(rm, 2, min)
plot(best.p,ty="l", col="green", main="k=0.6",ylab="Rating",xlab="Match",ylim=c(-3,3))
lines(worst.p, col="red")

rm <- tournament(n=10000, k=1, tv, cv)
best.p <- apply(rm, 2, max)
worst.p <- apply(rm, 2, min)
plot(best.p,ty="l", col="green", main="k=1",ylab="Rating",xlab="Match",ylim=c(-3,3))
lines(worst.p, col="red")
```

The smallest k values provide more stable results that are closer to the true ratings. The larger k values have much higher variance.

## Also : if we wanted to track the best and worst *players* over time:

```{r }
rm <- tournament(n=10000, k=0.01, tv, cv)
best.p <- apply(rm, 2, which.max)
worst.p <- apply(rm, 2, which.min)
plot(best.p,ty="p", col="green", main="k=0.01",ylab="Player",xlab="Match",ylim=c(0,13))
points(worst.p, col="red")
```

