---
title: "HovanesKeseyan_Homework5"
author: "Hovanes Keseyan"
date: "October 26, 2016"
output: html_document
---

## Part I - Optimization and standard errors

## 1 : Maximizing likelihood of gamma distribution on cats' hearts

```{r }
library(MASS)
data(cats)
summary(cats)

m <- mean(cats$Hwt)
v <- var(cats$Hwt)
a <- m*(((m*(1-m))/v)-1)
b <- (1-m)*(((m*(1-m))/v)-1)

ll <- function(p = c(a,b)) {
  ll.sum <- -sum(dbeta(cats$Hwt, p[1], p[2], log = TRUE))
  return(ll.sum)
}

nlm(ll, c(a,b))
```

THe nlm command is used to find the minimum negative of the log-likelihood, whose absolute value is equivalent to the maximum log-likelihood. We input the alpha and beta parameters from the gamma function that we obtained from the method of moments.

The estimate for alpha is -194.2401, while the estimate for beta is 175.9682. The log-likelihood there is 1.797693e+308. THe gradient is 0.

## 2 : make.gamma.loglike

```{r }
make.gamma.loglike <- function(p = c(a,b)) {
  loglike <- dbeta(cats$Hwt, p[1], p[2], log = TRUE)
  return(loglike)
}
```

## 3 : gamma.mle

```{r }
gamma.mle <- function()
```

## 

```{r }

```

## 

```{r }

```

## 

```{r }

```

## 

```{r }

```

## 

```{r }

```

## 

```{r }

```

