---
title: "HovanesKeseyan_Lab2"
author: "Hovanes Keseyan"
date: "October 4, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part I : Data Frames

```{r}
library(MASS)
data(Cars93)
View(Cars93)
```

## 1 : Inspecting the data structure

```{r }
summary(Cars93)
# ask for the number of rows in Cars93
nrow(Cars93)
```

By adding up the values in any of a number of columns of data (manufacturer, model, drivetrain, cylinders, make, etc) given by the summary() command, we can witness the number of rows in the data frame (93). The nrow() function also gives us the number of rows.

## 2 : Finding the mean of a subset of the data frame

```{r }
# create a new data frame rwd using only the rows from Cars93 that have the value "Rear" in the DriveTrain column
rwd <- Cars93[Cars93$DriveTrain=="Rear", ]
# display the DriveTrain column of rwd so we confirm we have pulled only the rows that correspond to rear-wheel drive cars
rwd$DriveTrain
# display the mean value of the column Price in the rwd data frame
mean(rwd$Price)
```

The mean proce of a car with rear-wheel drive is 28.95.

## 3 : Minimum value of a subset of the data frame

```{r }
# data frame of only rows containing cars with passengers of capacity 7
psg.7 <- Cars93[Cars93$Passengers == 7, ]
# data frame of only rows containing cars with passengers of capacity 6 or more
psg.6 <- Cars93[Cars93$Passengers >= 6, ]

# minimum horsepower of cars capacity 7
min(psg.7$Horsepower)
# minimum horsepower of cars capacity 6 or more
min(psg.6$Horsepower)
```

## 4 : Calculating travel distance

```{r }
# create an array of distance data that calculates distance by multiplying MPG by fuel capacity
hw.dist <- Cars93$MPG.highway * Cars93$Fuel.tank.capacity
# create a data frame that includes the new column of distance data
Cars93.dist <- data.frame(Cars93, hw.dist)
# display the row of data for the car with the maximum highway travel distance
Cars93.dist[Cars93.dist$hw.dist == max(Cars93.dist$hw.dist), ]
# display the row of data for the car with the minimum highway travel distance
Cars93.dist[Cars93.dist$hw.dist == min(Cars93.dist$hw.dist), ]
# display the row of data for the car with the median highway travel distance
Cars93.dist[Cars93.dist$hw.dist == median(Cars93.dist$hw.dist), ]
```

The two columns required to calculate travel distance are MPG highway and fuel tank capacity, because those two number multiplied will give the number of miles that the car can travel on the highway for one full tank.

## Part II : Reproducibility and Functions

```{r }
factory.function <- function (cars.output=1, trucks.output=1) {
  factory <- matrix(c(40,1,60,3),nrow=2,
    dimnames=list(c("labor","steel"),c("cars","trucks")))
  available <- c(1600,70); names(available) <- rownames(factory)
  slack <- c(8,1); names(slack) <- rownames(factory)
  output <- c(cars.output, trucks.output); names(output) <- colnames(factory)

  passes <- 0 # How many times have we  been around the loop?
  repeat {
     passes <- passes + 1
     needed <- factory %*% output # What do we need for that output level?
     # If we're not using too much, and are within the slack, we're done
     if (all(needed <= available) &&
         all((available - needed) <= slack)) {
       break()
     }
     # If we're using too much of everything, cut back by 10%
     if (all(needed > available)) {
       output <- output * 0.9
       next()
     }
     # If we're using too little of everything, increase by 10%
     if (all(needed < available)) {
       output <- output * 1.1
       next()
     }
     # If we're using too much of some resources but not others, randomly
     # tweak the plan by up to 10%
      # runif == Random number, UNIFormly distributed, not "run if"
     output <- output * (1+runif(length(output),min=-0.1,max=0.1))
  }
  
  return(output)
}
```

## 5 : Obtain output value

```{r }
factory.function()
```

The final output capacity is displayed, showing the output capacity for cars and trucks, respectively, for the factory.

## 6 : Repeating the function

```{r }
factory.function()
factory.function()
factory.function()
factory.function()
```

The output values are similar, but not identical. The runif() function randomly tweaks the plan, so we cannot expect these output values to always be equal.

## 7 : Return number of passes

```{r }
factory.function <- function (cars.output=1, trucks.output=1) {
  factory <- matrix(c(40,1,60,3),nrow=2,
    dimnames=list(c("labor","steel"),c("cars","trucks")))
  available <- c(1600,70); names(available) <- rownames(factory)
  slack <- c(8,1); names(slack) <- rownames(factory)
  output <- c(cars.output, trucks.output); names(output) <- colnames(factory)

  passes <- 0 # How many times have we  been around the loop?
  repeat {
     passes <- passes + 1
     needed <- factory %*% output # What do we need for that output level?
     # If we're not using too much, and are within the slack, we're done
     if (all(needed <= available) &&
         all((available - needed) <= slack)) {
       break()
     }
     # If we're using too much of everything, cut back by 10%
     if (all(needed > available)) {
       output <- output * 0.9
       next()
     }
     # If we're using too little of everything, increase by 10%
     if (all(needed < available)) {
       output <- output * 1.1
       next()
     }
     # If we're using too much of some resources but not others, randomly
     # tweak the plan by up to 10%
      # runif == Random number, UNIFormly distributed, not "run if"
     output <- output * (1+runif(length(output),min=-0.1,max=0.1))
     }
  output.matrix <- c(output, passes)    # add the number of passes to the output array
  names(output.matrix) <- c("cars", "trucks", "passes")  # name the columns
  return(output.matrix)
}

factory.function()

```

## 8 : Changing initial output and returning final demand

```{r }
factory.function <- function (cars.output=30, trucks.output=20) {
  factory <- matrix(c(40,1,60,3),nrow=2,
    dimnames=list(c("labor","steel"),c("cars","trucks")))
  available <- c(1600,70); names(available) <- rownames(factory)
  slack <- c(8,1); names(slack) <- rownames(factory)
  output <- c(cars.output, trucks.output); names(output) <- colnames(factory)

  passes <- 0 # How many times have we  been around the loop?
  repeat {
     passes <- passes + 1
     needed <- factory %*% output # What do we need for that output level?
     # If we're not using too much, and are within the slack, we're done
     if (all(needed <= available) &&
         all((available - needed) <= slack)) {
       break()
     }
     # If we're using too much of everything, cut back by 10%
     if (all(needed > available)) {
       output <- output * 0.9
       next()
     }
     # If we're using too little of everything, increase by 10%
     if (all(needed < available)) {
       output <- output * 1.1
       next()
     }
     # If we're using too much of some resources but not others, randomly
     # tweak the plan by up to 10%
      # runif == Random number, UNIFormly distributed, not "run if"
     output <- output * (1+runif(length(output),min=-0.1,max=0.1))
     }
  output.matrix <- c(output, needed, passes)    # add the number of passes to the output array
  names(output.matrix) <- c("cars", "trucks", "labor.needed", "steel.needed" , "passes")  # name the columns
  return(output.matrix)
}

factory.function()
```

The final output plan is similar to what it was before. The final demand for resources is listed in the output for lebor and steel, respectively. THe plan is within budget and within the slack, since the labor is within 8 of and under 1600, and steel is within 1 of and under 70. The number of iterations is also listed in the output indicating passes.