---
title: "HovanesKeseyan_Lab9"
author: "Hovanes Keseyan"
date: "November 22, 2016"
output: pdf_document
---

## Gibbs sampler

## 1 : Full conditional distributions of $Y_1|Y_2$ and $Y_2|Y_1$

```{r }
x <- c(1, 1, 1, 1, 1, 1, 1, 1)
y1.y2 <- function(x, p, y2){
  pnorm(x, p*y2, 1-p^2)
}
y2.y1 <- function(x, p, y1){
  pnorm(x, p*y1, 1-p^2)
}
y1.y2(x, .5, 1)
y2.y1(x, .5, 1)
```
$Y_1|Y_2 = N(\mu_1 + \Sigma_{12}\Sigma_{22}^{-1}(Y_2-\mu_2), \Sigma_{11}-\Sigma_{12}\Sigma_{11}^{-1}\Sigma_{21})$

$Y_1|Y_2 = N(pY_2, 1-p^2)$


$Y_2|Y_1 = N(\mu_2 + \Sigma_{21}\Sigma_{11}^{-1}(Y_1-\mu_1), \Sigma_{22}-\Sigma_{21}\Sigma_{22}^{-1}\Sigma_{12})$

$Y_2|Y_1 = N(pY_1, 1-p^2)$

## 2 : Gibbs sampler

```{r }
gibbs.chain <- function(n, y1.start = 0, y2.start = 0, p) {
    output <- matrix(0, nrow=n, ncol=2)
    for(i in 1:n){
        y1.new <- rnorm(1, p*y2.start, 1-p^2)
        y2.new <- rnorm(1, y1.new, 1-p^2)
        output[i,] <- c(y1.new, y2.new)
        y1.start <- y1.new
        y2.start <- y2.new    
    }
    return(output)
}
```


## 3 : Generating 10000 draws

```{r }
samp <- gibbs.chain(10000, 0, 0, 0.7)
y1 <- samp[,1]
y2 <- samp[,2]

plot(y1, type = "l")
plot(y2, type = "l")
```


## 4 : PLotting marginal distributions with true distribution

```{r }
hist(y1, probability = TRUE)
curve(dnorm(x), add = TRUE)

hist(y2, probability = TRUE)
curve(dnorm(x), add = TRUE)
```

The marginal distributions for Y1 and Y2 seem to follow the true distribution very well by shape.

## 5 : Estimating effective sample size

```{r }
library(mcmcse)
ess(y1)
ess(y2)

par(mfrow=c(1,2))
estvssamp(y1)
estvssamp(y2)
```

We could have used a much smaller sample size than the original 10000 that we used, and we would have gotten results that matched the true distribution while using less computing power. 

## 6 : Mixing properties for the Gibbs sampler

The mixing properties for the Gibbs sampler allow for a robust approximation of the distribution. The histogram plots of the marginal distributions with the true normal distributions overlayed support this.

